---
title: "R sessions"
author: "Eric Macwan"
date: "27/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The R programming can be carried out as interative R-session. To start an R sisson, type 'R' from the command line in windows or linux OS. For example, from shell prompt '$' in lunux

`$ R`

This generates the following output before entering  '>' prompt of R:

```
R version 3.4.4 (2018-03-15) -- "Someone to Lean On"
Copyright (C) 2018 The R Foundation for Statistical Computing
Platform: x86_64-pc-linux-gnu (64-bit)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type 'license()' or 'licence()' for distribution details.

  Natural language support but running in an English locale

R is a collaborative project with many contributors.
Type 'contributors()' for more information and
'citation()' on how to cite R or R packages in publications.

Type 'demo()' for some demos, 'help()' for on-line help, or
'help.start()' for an HTML browser interface to help.
Type 'q()' to quit R.

[Previously saved workspace restored]

>
```

Once we are inside an R session, we can directly execute R language commands by typing them line by line. Pressing the *return key* terminates typing of command and brings the > prompt again. In the example session below, we declare 2 variables 'a' and 'b' to have values 5 and 6 respectively, and assign their sum to another variable called 'res':

```{r}
a = 5
b = 6
c = a+b
c
```


#### **To get help on any function of R**

Type help(function-name) in R prompt. For example, if we need help on "if" logic, type,

```{r}
help("if")
```

then, manual of asked function is printed.


#### **To exit the R session**

Type quit() in R prompt, and say 'n' (no) for saving the workspace image. This means, we do not want to save the memory of all the commands we typed in the current session:

`>quit()`

`save workspace image? [y/n/c]: n `


#### **Saving the R session** 

Note that by not saving the current session, we loose all the memory of current session commands and the variables and objects created.
  
When we work in R, the R objects we created and loaded are stored in a memory portion called **workspace**. When we say 'no' to saving the workspace, all these objects are wiped out from the workspace memory. With 'yes', they are saved into a file called ".RData" is written to the present "working directory".
  
In Linux, this "working directory" is generally the directory from where the R was stared through command 'R'. In windows, it can be either "My Documents" or user's home directory.

When we start R in the same crrent directory next time, the workspace and all the created objects are restored automatically from this ".RData" directory.

#### **Listing the objects in the current R session**

We can list the names of the objects in the current R session by `ls()` command. For example, start R session fresh and proceed as follows:

```{r}
a = 5
b = 6
c = 8
sum = a+b+c

sum
```
```{r}
ls()
```


Here, Variable names created in current session is get printed.

#### **Removing objects from the current R session**

Specific object created in the current session can be removed using `rm()` command. If we specify the name of an object, it will be removed. If we just say `rm(list = las())`, all objects created so far will be removed. Example:

```{r}
a = 5
b = 6
c = 8
sum = a+b+c

sum
```
```{r}
ls()
```
```{r}
rm(list = c("sum"))

ls()
```
```{r}
rm(list = ls())

ls()
```

#### **Getting and setting the current working directories**

From R prompt, we can get information about the current working directory by `getwd()` command:

```{r}
getwd()
```
Similarly, we can set the current work directory by calling `setwd()` function:
```{r}
setwd("/home/eric/Documents/projects/")
```

```{r include=FALSE}
setwd("/media/eric/nebula/projects/R_Fundamentals")
```

