---
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In R, we can assign integer and floating point values to variables directly. The mathematical operations can be performed with symbols following a format exactly similar to languages like C, C++, java, python and Perl. We have already performed a simple operation `c = a+b` in the previous section.

For example, we can perform the operation $$c = \frac{a+b}{a*b}$$ by assigning values directly at declaration.

```{r}
a = 7.5
b = 6
c = (a+b)/(a*b)
c
```
In another example, we apply the formula $$ y = \sqrt{1.0+(z/r)^0.38} $$ as,
```{r}
z = 22.9
r = 6.7
y = sqrt(1.0+(z/r)^0.38)
y
```
Note the explicit usage of brackets for grouping the term to remove ambiguity

A lost of useful inbuilt functions in R is given below:

```
Functions                   Description
abs(x)----------------------absolute vlaue
sqrt(x)---------------------square root
ceiling(x)------------------ceiling(3.475) is 4
floor(x)--------------------floor(3.475) is 3
trunc(x)--------------------trunc(5.99) is 5
round(x, digits=n)----------round(3.457, digits=2) is 3.48
signif(x, digits=n)---------signif(3.457, digits=2) is 3.5
cos(x),sin(x),tan(x)--------Triginometric sine, cosine and tan functions
acos(x),cosh(x),acosh(x)----arcsine, arccosine and arctangent functions
log(x)----------------------natural logarithm
log10(x)--------------------common logarithm
exp(x)----------------------e^{x}
```



